---
title: "risk set matching analysis "
output: 
  html_document:
  code_folding: hide
editor_options: 
  chunk_output_type: console
---

# Setup
Load packages
```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(knitr)
library(rsmatch)
library(survival)
library(nbpMatching)
library(MatchIt)
```

Load files
```{r}
setwd("/Users/emma/Desktop/local_only/mind_study")

# adt_long <- read_csv("./adt_long.csv")
# adt_long_patients <- read_csv("./adt_long_patients.csv")
# covid_long <- read_csv("./covid_long.csv")
# micro_long <- read_csv("./micro_long.csv")
# micro_long_positives_only <- read_csv("./micro_long_positives_only.csv")
# patients <- read_csv("./patients.csv", col_types = cols(hospice_dt = col_datetime(format = "")))
# bacterial_infections_blood <- read_csv("./bacterial_infections_blood.csv")
# bacterial_infections_urine <- read_csv("./bacterial_infections_urine.csv")
# bacterial_infections_resp <- read_csv("./bacterial_infections_resp.csv")
covid_admissions_bact_blood <- read_csv("./covid_admissions_bact_blood.csv")
covid_admissions_bact_urine <- read_csv("./covid_admissions_bact_urine.csv")
covid_admissions_bact_resp <- read_csv("./covid_admissions_bact_resp.csv")
```

Set options to vary analysis: uncomment section we want to run.
```{r}
## blood infections:
# df <- covid_admissions_bact_blood
# unmatchedplot <- "LOS unmatched - blood"
# plot1title <- "LOS after matching - blood"
# plot2title <- "LOS after matching, only uninfected controls - blood"
# tabletitle <- "Risk-set matching results - blood"

# urine infections:
df <- covid_admissions_bact_urine
unmatchedplot <- "LOS unmatched - urine"
plot1title <- "LOS after matching - urine"
plot2title <- "LOS after matching, only uninfected controls - urine"
tabletitle <- "Risk-set matching results - urine"
# 
# resp infections:
# df <- covid_admissions_bact_resp
# unmatchedplot <- "LOS unmatched - respiratory"
# plot1title <- "LOS after matching - respiratory"
# plot2title <- "LOS after matching, only uninfected controls - respiratory"
# tabletitle <- "Risk-set matching results - respiratory"
```

Options for individual pathogens:
```{r}
# Top 5 blood pathogens:
#path = "KLEBSIELLA PNEUMONIAE"
#path = "METHICILLIN-SUSCEPTIBLE STAPHYLOCOCCUS AUREUS"
#path = "ESCHERICHIA COLI"
#path = "PSEUDOMONAS AERUGINOSA"
#path = "ENTEROCOCCUS FAECALIS"

# top 5 urine pathogens:
 path = "ESCHERICHIA COLI"
# path = "CANDIDA ALBICANS"
# path = "KLEBSIELLA PNEUMONIAE"
# path = "CANDIDA GLABRATA"
# path = "ENTEROCOCCUS FAECALIS"

# top 5 respiratory pathogens:
# path = "PSEUDOMONAS AERUGINOSA"
# path = "KLEBSIELLA PNEUMONIAE"
# path = "METHICILLIN-SUSCEPTIBLE STAPHYLOCOCCUS AUREUS"
# path = "METHICILLIN-RESISTANT STAPHYLOCOCCUS AUREUS"
# path = "KLEBSIELLA (FORMERLY ENTEROBACTER) AEROGENES"
```


# Exploratory code also in mind_exploratory rmd file
Most of the following code is also in the mind_exploratory rmd file. Revisit questions about how many covid admissions each patient has and how many bacterial infections each patient has occuring during covid admissions:
```{r}
# # how many "covid admissions" does each patient have?
# adt_long_patients %>% 
#   group_by(mrn) %>% 
#   summarize(n_covid_admissions = sum(covid_admission)) %>% 
#   ungroup() %>% 
#   group_by(n_covid_admissions) %>% 
#   summarize(n = n()) %>% 
#   kable()
# 
# # How many infections did each covid patient have that occurred during covid inpatient admissions?
# covid_admissions_bacterial %>% 
#   group_by(mrn) %>% 
#   summarize(n_bacterial_infections = sum(bacterial_infection)) %>% 
#   ungroup() %>%  group_by(n_bacterial_infections) %>% 
#   summarize(n = n()) %>% 
#   kable()
#     # >1 means multiple infections 
# 
# # How many unique mrns are there?
# covid_admissions_bacterial %>%  count(mrn)
# 
# # examine details of covid admissions for patients with multiple covid admissions:
# adt_long_patients %>% 
#   group_by(mrn) %>% 
#   mutate(n_covid_admissions = sum(covid_admission)) %>% 
#   filter(n_covid_admissions > 1,
#          covid_admission == T) %>% 
#   select(mrn, encounter_id, adt_inpatient_dt, adt_discharge_dt, adt_site, adt_care_level, adt_discharge_disp) %>% 
#   view()
```
It seems that these multiple inpatients stays either have to do with transfer to a different care level (ICU, hospice, etc), 
or with multiple stays that are initiated by emergency room visits with the week following covid diagnosis.
I may want to look for transfers to ICU care that happen later in the trajectory for patients with covid admissions - 
ie, if they were admitted and then transferred and the transfer shows up as a different line.

# Plot length of stay pre-matching

```{r}
df %>% 
  group_by(mrn) %>% 
  summarize(first_inpatient_dt = min(adt_inpatient_dt),
            last_discharge_dt = max(adt_discharge_dt),
            bacterial_infection = as.logical(max(bacterial_infection))) %>% 
  mutate(covid_time_in_hospital = last_discharge_dt - first_inpatient_dt,
         covid_days_in_hospital = covid_time_in_hospital / ddays(1)) %>%
  ungroup() %>% 
  ggplot(aes(x = covid_days_in_hospital, color = bacterial_infection)) +
  geom_density() +
  scale_color_manual(values = alpha(c("red", "blue"))) +
  labs(title = unmatchedplot, 
       x = "Length of stay (days)")
```

ALTERNATE FOR SINGLE PATHOGEN ANALYSIS:
```{r}
df %>% 
  mutate(path_of_interest = case_when(organism_name == path ~ T,
                                      organism_name != path ~ F,
                                      is.na(organism_name) ~ F)) %>% 
  group_by(mrn) %>% 
  summarize(first_inpatient_dt = min(adt_inpatient_dt),
            last_discharge_dt = max(adt_discharge_dt),
            bacterial_infection = as.logical(max(bacterial_infection)),
            path_of_interest = as.logical(max(path_of_interest))) %>% 
  mutate(covid_time_in_hospital = last_discharge_dt - first_inpatient_dt,
         covid_days_in_hospital = covid_time_in_hospital / ddays(1)) %>%
  ungroup() %>% 
  ggplot(aes(x = covid_days_in_hospital, color = path_of_interest)) +
  geom_density() +
  scale_color_manual(values = alpha(c("red", "blue"))) +
  labs(title = paste(unmatchedplot, "-", path), 
       x = "Length of stay (days)")
```

# Initial steps to set up data for matching:

First step of setting up data for matching:
8/6 - to do: update some of these to match "working with dates" instructions from https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html#Part_1:_Introduction_to_Survival_Analysis
```{r}
matching_df_all_vars <- df %>% 
  group_by(mrn) %>% 
  mutate(first_inpatient_dt = min(adt_inpatient_dt),
         last_discharge_dt = max(adt_discharge_dt),
         covid_time_in_hospital = last_discharge_dt - first_inpatient_dt,
         covid_days_in_hospital = covid_time_in_hospital / ddays(1),
         first_bacterial_start_dt = min(first_dt, na.rm = T),
         age_at_admission = as.numeric((as_date(first_inpatient_dt) - as_date(dob))/365),
         bacterial_infection_ever = replace_na(bacterial_infection_ever, 0),
         bacterial_infection_ever = as.logical(bacterial_infection_ever),
         covid_and_bacterial_same_admission = replace_na(covid_and_bacterial_same_admission, F),
         combined_race_ethnicity = factor(combined_race_ethnicity, 
                                          levels = c("White", "BLACK OR AFRICAN AMERICAN", "Hispanic/Latino", "other")),
         sex = factor(sex, levels = c("male", "female")),
         trt_time = case_when(!(first_bacterial_start_dt == Inf) ~ as.numeric(((first_bacterial_start_dt) - (first_inpatient_dt)) / ddays(1)))) %>% 
  group_by(mrn) %>% arrange(mrn, first_dt) %>% mutate(index = row_number()) %>% 
    # index sets it up for us to choose just one row per mrn (with info for the first infection)
    # This gets rid of duplicate rows per mrn because of duplicate admissions or duplicate bacterial infections
    # Somewhat similar to just using the summarize command, but I wanted to preserve some of the variables that are 
    # available on every row without having to include them explicitly in summarize, so I did it this way.
  filter(index == 1)
```

INDIVIDUAL PATHOGEN VERSION:

```{r}

# Step 1: set up dataframe with just MRN, time-invarying covariates, time of censoring or event, and event 1 or 0 (where event is onset of infection with path of interest)

path_of_interest <- df %>% 
  group_by(mrn) %>% 
  mutate(first_inpatient_dt = as_datetime(min(adt_inpatient_dt)), # beginning of hospital stay
         last_discharge_dt = as_datetime(max(adt_discharge_dt)), # end of hospital stay
         age_at_admission = as.numeric((as_date(first_inpatient_dt) - as_date(dob))/365.25),
         combined_race_ethnicity = factor(combined_race_ethnicity, 
                                          levels = c("White", "BLACK OR AFRICAN AMERICAN", "Hispanic/Latino", "other")),
         sex = factor(sex, levels = c("male", "female")),
         path_of_interest_dt = ifelse(organism_name == path, first_dt, NA),
         # path_of_interest_dt: date of detection of infection with the pathogen of interest, na if not pathogen of interest
         first_path_of_interest_dt = ifelse(all(is.na(path_of_interest_dt)), NA, as_datetime(min(path_of_interest_dt, na.rm = T))),
         # first_path_of_interest_dt: date of first detection of path of interest if patient ever had pathogen of interest
         test1 = last_discharge_dt - first_inpatient_dt,
         test2 = first_path_of_interest_dt - first_inpatient_dt,
         #test2 = ifelse(is.na(first_path_of_interest_dt), NA, first_path_of_interest_dt - first_inpatient_dt),
        # time = ifelse(is.na(first_path_of_interest_dt),
         #              as.numeric(((last_discharge_dt) - (first_inpatient_dt)) / ddays(1)),
         #              as.numeric(((first_path_of_interest_dt) - (first_inpatient_dt)) / ddays(1))),
         # time: time until event if patient had one (detection of pathogen of interest) OR censoring (discharge from hospital)
         event = !is.na(path_of_interest_dt)) %>% 
  

  ## 1/18/22: something's not working, that's why i made "test" vars, i dunno, fuck it
  
  arrange(mrn, first_dt) %>% mutate(index = row_number()) %>% filter(index == 1) %>% 
    # index sets it up for us to choose just one row per mrn, which should include all information we need.
    # This gets rid of duplicate rows per mrn because of duplicate admissions or duplicate bacterial infections
    # Somewhat similar to just using the summarize command, but I wanted to preserve some of the variables that are 
    # available on every row without having to include them explicitly in summarize, so I did it this way.
  select(mrn, first_inpatient_dt, last_discharge_dt, age_at_admission, combined_race_ethnicity, sex, first_path_of_interest_dt, time, event)

# Step 2: set up dataframe with one row per MRN, time = 0 and other infection = 0

# Step 3: set up dataframe with just MRN, time of onset of other infection (that's not path of interest) and other_infection = 1 (since we're only keeping lines that do have another infection with a different pathogen)

# Step 4: combine 2 and 3

# step 5: make a similar thing but with ICU stays

# step 6: use tmerge on 1,4 and 5 to set up data for coxph. 
  

# step 7???? might need to set up the rest of the matching df (like, time in hospital, etc) to use for the matching step later. 
```


## Results DF
Set up dataframe to record results:
```{r}
ttest_output <- tibble(sample = c("unmatched", "matched", "matched - uninfected controls only"), 
                       estimate = c(0, 0, 0),
                       CI_lb = c(0, 0, 0),
                       CI_ub = c(0,0,0))
```

# Matching:

I'll run the coxph first, then loop through a simple matching algorithm

## Setup for coxph. 

Check for missingness, since we will get rid of anything with missing variables:
```{r}
for (v in c("age_at_admission", "sex", "adt_status", "covid_visits_start_date", "combined_race_ethnicity", "insurance_type", "bmi", "dx_ckd", "dx_dm", "dx_htn",  "dx_cad",  "dx_hrt_failure")) {
  n_missing <- matching_df_all_vars %>% 
    filter(is.na(v)) %>% 
    summarize(n = n()) %>% 
    unlist()
  print(n_missing)
}
matching_df_all_vars %>% ungroup() %>% filter(is.na(age_at_admission)) %>%  summarize(n = n())
matching_df_all_vars %>% ungroup() %>% filter(is.na(sex)) %>%  summarize(n = n())
matching_df_all_vars %>% ungroup() %>% filter(is.na(adt_status)) %>%  summarize(n = n())
matching_df_all_vars %>% ungroup() %>% filter(is.na(covid_visits_start_date)) %>%  summarize(n = n())
matching_df_all_vars %>% ungroup() %>% filter(is.na(combined_race_ethnicity)) %>%  summarize(n = n())
matching_df_all_vars %>% ungroup() %>% filter(is.na(insurance_type)) %>%  summarize(n = n()) #missing 333
matching_df_all_vars %>% ungroup() %>% filter(is.na(bmi)) %>%  summarize(n = n()) # - many bmi missing missing 589
matching_df_all_vars %>% ungroup() %>% filter(is.na(dx_ckd)) %>%  summarize(n = n())
matching_df_all_vars %>% ungroup() %>% filter(is.na(dx_dm)) %>%  summarize(n = n())
matching_df_all_vars %>% ungroup() %>% filter(is.na(dx_htn)) %>%  summarize(n = n())
matching_df_all_vars %>% ungroup() %>% filter(is.na(dx_cad)) %>%  summarize(n = n())
matching_df_all_vars %>% ungroup() %>% filter(is.na(dx_hrt_failure)) %>%  summarize(n = n())

```


In this code chunk, we add a time variable and exclude cases that have any missing values. This may need to be amended later, but should work for the moment.(9/10/21)
```{r}
# coxph_df <- matching_df_all_vars %>% 
#   mutate(time = case_when(is.na(trt_time) == F ~ trt_time, 
#                           #time is time of bacterial infection (trt_time) if there was one
#                           is.na(trt_time) == T ~ covid_days_in_hospital),
#                           # time is censoring time (days in hospital) if there wasn't an infection
#          status = !is.na(trt_time)) %>% 
#         # status is F if no infection (and therefore no "treatment time") and T if there was an infection
#   filter(any(is.na(age_at_admission), is.na(sex), is.na(covid_visits_start_date), is.na(combined_race_ethnicity), is.na(bmi), is.na(dx_ckd), is.na(dx_dm), is.na(dx_htn), is.na(dx_cad), is.na(dx_hrt_failure), is.na(time), is.na(status)) == F) 
#   # filter out anything with a missing value for any of our predictor variables or status or time value
```

ALTERNATE COXPH_DF: USE FOR INDIVIDUAL PATHOGEN ANALYSES.
COMMENT OUT SECTION ABOVE.
What's different about this version: variable "time" will be used to do the cox analysis, and refers to the time of infection of the pathogen of interest ONLY and otherwise is the time in hospital. variable "time_any_pathogen" is the time of infection for anyone who has had an infection at the site with any pathogen. It will be used for the sole purpose of determining the risk set from which to draw controls (ie, not including individuals who have already been infected with a different pathogen already as an eligible control.)
```{r}
coxph_df <- matching_df_all_vars %>% 
  mutate(time = case_when(is.na(trt_time) == F & organism_name == path ~ trt_time, 
                          #time is time of bacterial infection of the pathogen of interest (trt_time) if there was one
                          is.na(trt_time) == T ~ covid_days_in_hospital,
                          organism_name != path ~ covid_days_in_hospital),
                          # time is censoring time (days in hospital) if there wasn't an infection of the pathogen of interest.
         time_any_pathogen = case_when(is.na(trt_time) == F ~ trt_time, 
                          #time is time of bacterial infection (trt_time) if there was one
                          is.na(trt_time) == T ~ covid_days_in_hospital),
                          # time is censoring time (days in hospital) if there wasn't an infection
         status = (organism_name == path)) %>% 
        # status is F if no infection (and therefore no "treatment time") and T if there was an infection
  filter(any(is.na(age_at_admission), is.na(sex), is.na(covid_visits_start_date), is.na(combined_race_ethnicity), is.na(bmi), is.na(dx_ckd), is.na(dx_dm), is.na(dx_htn), is.na(dx_cad), is.na(dx_hrt_failure), is.na(time), is.na(status)) == F) 
  # filter out anything with a missing value for any of our predictor variables or status or time value
```


Checkpoint: did I correctly set status to be T for people who had infections, F otherwise?
```{r}
# coxph_df %>% 
#   ungroup() %>% 
#   summarize(prop_status_1 = mean(status),
#             n_with_status_1 = sum(status),
#             prop_covid_bacterial_same_admission = mean(covid_and_bacterial_same_admission),
#             n_covid_bacterial_same_admission = sum(covid_and_bacterial_same_admission))
```
9/10/21: These numbers should be the same, and yet they are slightly different! Need to look into why. Why do we have status = 1 for some people who are NOT listed as having covid and bacterial infections in the same admission? Look into how each variable was created to figure out. Not a big impediment, but need to figure out.
11/22/21: probably don't worry about this, now that I'm using this file for multiple types of infections and covid_bacterial_same_admission is only for bloodstream infections, I think.

## Set up cox model:
```{r}
coxph_model <- coxph(Surv(time, status) ~ age_at_admission + sex + covid_visits_start_date + combined_race_ethnicity + bmi + dx_dm + dx_htn + dx_cad + dx_hrt_failure,
               data = coxph_df)

coxph_model %>% 
  gtsummary::tbl_regression(exp = TRUE, estimate_fun = function(x) style_ratio(x, digits = 4))
# Of course I should add quite a few other variables here - however, I think for the moment, I'll leave like this just to make sure I can set this up to run properly. 

propensity <- predict(coxph_model, type = "risk", x = coxph_df)
```


9/1/21: notes on things still to do:
- adapt variables to have appropriate reference groups.
- why doesn't propensity come out as as long as the coxph_model dataframe? Are there a bunch missing all their variables or soemthing????
    Update 9/2: had to do with missing values. updating to napredict might work.....

## Matching procedure

Set up "risk set" dataframe, which will include everyone for starters. This will be revised each time matching happens.
```{r}
riskset <- tibble(coxph_df, propensity) %>% 
  select(mrn, status, propensity, time, time_any_pathogen, covid_days_in_hospital)
```

Set up dataframe of infection times with space for output:
```{r}
matches <- riskset %>% 
  filter(status == 1) %>% 
  select(mrn, propensity, time, time_any_pathogen, status, covid_days_in_hospital) %>% 
  arrange(time) %>% 
  mutate(ctrl_mrn = NA, ctrl_propensity = NA, ctrl_time = NA, ctrl_status = NA, ctrl_covid_days_in_hospital = NA)
```

Now run the matching process:
```{r}
# for(i in 1:nrow(matches)) {
#   # pick control among eligible riskset based on propensity
#   temp <- riskset %>% 
#     filter(time >= matches$time[i], # riskset only includes those still at risk
#            mrn != matches$mrn[i]) %>% # and not infection we're trying to match
#     mutate(propensity_difference = abs(propensity - matches$propensity[i])) %>% 
#     filter(propensity_difference == min(propensity_difference))
#       # keep control with closest risk to the infection we're matching
#   
#   # plug info about control into matches df
#   matches$ctrl_mrn[i] <- temp$mrn[1]
#   matches$ctrl_propensity[i] <- temp$propensity[1]
#   matches$ctrl_time[i] <- temp$time[1]
#   matches$ctrl_status[i] <- temp$status[1]
#   matches$ctrl_covid_days_in_hospital[i] <- temp$covid_days_in_hospital[1]
#   
#   # save new riskset excluding the control that was just matched 
#   riskset <- riskset %>% filter(mrn!=temp$mrn[1])
# }
```

SINGLE PATHOGEN VARIATION: time_any_path must be greater than the next exposed person's time_any_path - ie, this is where we exclude people already infected with a different pathogen from the potential control group.
```{r}
for(i in 1:nrow(matches)) {
  # pick control among eligible riskset based on propensity
  temp <- riskset %>% 
    filter(time_any_pathogen >= matches$time_any_pathogen[i], # riskset only includes those still at risk
           mrn != matches$mrn[i]) %>% # and not infection we're trying to match
    mutate(propensity_difference = abs(propensity - matches$propensity[i])) %>% 
    filter(propensity_difference == min(propensity_difference))
      # keep control with closest risk to the infection we're matching
  
  # plug info about control into matches df
  matches$ctrl_mrn[i] <- temp$mrn[1]
  matches$ctrl_propensity[i] <- temp$propensity[1]
  matches$ctrl_time[i] <- temp$time[1]
  matches$ctrl_status[i] <- temp$status[1]
  matches$ctrl_covid_days_in_hospital[i] <- temp$covid_days_in_hospital[1]
  
  # save new riskset excluding the control that was just matched 
  riskset <- riskset %>% filter(mrn!=temp$mrn[1])
}
```

For now, we've included infected people as potential controls if they're still in the risk set, but this could be easily changed. 
Assess how often people with infections got selected as controls:
```{r}
matches %>% 
  summarize(n_infected_controls = sum(ctrl_status))
```
 This is higher than would be seen at random, but is consistent with literature (at least in that "later exposed" ppl are selected as controls more often than random), because presumably these individuals have commonalities with the infected individuals they serve as controls for.

# Make some comparisons in the datasets and between matches:

Set up dataset with all data and the matched individuals only
```{r}
# make a list of ids of the people who were matched
matched_ppl <- c(unlist(matches$mrn), unlist(matches$ctrl_mrn))

# make dataset of matching ppl only
matched_ppl_df <- coxph_df %>% filter(mrn %in% matched_ppl)
```

Check covariate balance:
```{r}
# # compare baseline characteristics in the unmatched sample:
# example_df %>% 
#   group_by(status) %>% 
#   summarize(mean_cov1 = mean(cov1),
#             sd_cov1 = sd(cov1),
#             mean_cov2 = mean(cov2),
#             sd_cov2 = sd(cov2),
#             mean_cov3 = mean(cov3),
#             sd_cov3 = sd(cov3))
# 
# # compare baseline characteristics in the matched sample:
# matched_ppl_df %>% 
#   group_by(status) %>% 
#   summarize(mean_cov1 = mean(cov1),
#             sd_cov1 = sd(cov1),
#             mean_cov2 = mean(cov2),
#             sd_cov2 = sd(cov2),
#             mean_cov3 = mean(cov3),
#             sd_cov3 = sd(cov3))
# 
# # following Hade et al, look at covariate impact on treatment hazard before and after matching:
# # before matching:
# coxph_model
# 
# #after matching:
# coxph_model_matched <- coxph(Surv(time, status) ~ cov1 + cov2 + cov3, data = matched_ppl_df)
# coxph_model_matched
```

Notes: more resources on checking balance:
https://cran.r-project.org/web/packages/MatchIt/vignettes/assessing-balance.html
https://onlinelibrary-wiley-com.ezproxy.cul.columbia.edu/doi/epdf/10.1002/sim.3697

In this very small example, there was already a pretty good balance between groups in the 
covariates before matchin, and comparable balance after matching. I don't expect this to be the 
case with the real data.

# Look at results:

```{r}
# make subsets of data to use for t-tests (there's surely a quicker way, but this should work)
infected <- matches %>% select(mrn:covid_days_in_hospital)
all_uninfected <- coxph_df %>% filter(status == F)
controls <- matches %>% select(ctrl_mrn:ctrl_covid_days_in_hospital)
  

# unmatched sample:
unmatched_ttest <-t.test(x = infected$covid_days_in_hospital, y = all_uninfected$covid_days_in_hospital) 

# matched sample:
#t.test(x = infected$covid_days_in_hospital, y = controls$ctrl_covid_days_in_hospital)

# matched paired sample t-test (following watson et al for this)
matched_ttest <- matches %>% mutate(diff = covid_days_in_hospital - ctrl_covid_days_in_hospital) %>% 
  pull(diff) %>% 
  t.test()

# put ttest results in table
ttest_output[1,2] <- unmatched_ttest$estimate["mean of x"] - unmatched_ttest$estimate["mean of y"]
ttest_output[1,3] <- unmatched_ttest$conf.int[1]
ttest_output[1,4] <- unmatched_ttest$conf.int[2]
ttest_output[2,2] <- matched_ttest$estimate
ttest_output[2,3] <- matched_ttest$conf.int[1]
ttest_output[2,4] <- matched_ttest$conf.int[2]
```

Plot results:
```{r}
matches %>% 
  ggplot(aes(x = covid_days_in_hospital)) +
  geom_density(color = "blue") +
  geom_density(aes(x = ctrl_covid_days_in_hospital), color = "red") +
  labs(title = paste(plot1title, "-", path), 
       x = "Length of stay (days)")
```

## Now, consider only never-infected people as controls:

Set up "risk set" dataframe, which will include ONLY UNINFECTED people
```{r}
riskset2 <- tibble(coxph_df, propensity) %>% 
  select(mrn, status, propensity, time, covid_days_in_hospital) %>% 
  filter(status == F)

infected2 <- tibble(coxph_df, propensity) %>% 
  select(mrn, status, propensity, time, covid_days_in_hospital) %>% 
  filter(status == T)
```

Set up dataframe of infection times with space for output:
```{r}
matches2 <- infected2 %>% 
  select(mrn, propensity, time, status, covid_days_in_hospital) %>% 
  arrange(time) %>% 
  mutate(ctrl_mrn = NA, ctrl_propensity = NA, ctrl_time = NA, ctrl_status = NA, ctrl_covid_days_in_hospital = NA)
```

Now run the matching process:
```{r}
for(i in 1:nrow(matches2)) {
  # pick control among eligible riskset based on propensity
  temp <- riskset2 %>% 
    filter(time >= matches2$time[i], # riskset only includes those still at risk
           mrn != matches2$mrn[i]) %>% # and not infection we're trying to match
    mutate(propensity_difference = abs(propensity - matches2$propensity[i])) %>% 
    filter(propensity_difference == min(propensity_difference))
      # keep control with closest risk to the infection we're matching
  
  # plug info about control into matches df
  matches2$ctrl_mrn[i] <- temp$mrn[1]
  matches2$ctrl_propensity[i] <- temp$propensity[1]
  matches2$ctrl_time[i] <- temp$time[1]
  matches2$ctrl_status[i] <- temp$status[1]
  matches2$ctrl_covid_days_in_hospital[i] <- temp$covid_days_in_hospital[1]
  
  # save new riskset excluding the control that was just matched 
  riskset2 <- riskset2 %>% filter(mrn!=temp$mrn[1])
}
```

## Look at results using matches2:

# Look at results:

```{r}
# make subsets of data to use for t-tests (there's surely a quicker way, but this should work)
# infected <- matches %>% select(mrn:covid_days_in_hospital)
# all_uninfected <- coxph_df %>% filter(status == F)
controls2 <- matches2 %>% select(ctrl_mrn:ctrl_covid_days_in_hospital) 

# unmatched sample:
t.test(x = infected$covid_days_in_hospital, y = all_uninfected$covid_days_in_hospital)

# matched sample:
#t.test(x = infected$covid_days_in_hospital, y = controls2$ctrl_covid_days_in_hospital)

# paired sample t-test (following watson et al for this)
matched_ttest2 <- matches2 %>% mutate(diff = covid_days_in_hospital - ctrl_covid_days_in_hospital) %>% 
  pull(diff) %>% # pick 
  t.test()

# put ttest results in df
ttest_output[3,2] <- matched_ttest2$estimate
ttest_output[3,3] <- matched_ttest2$conf.int[1]
ttest_output[3,4] <- matched_ttest2$conf.int[2]
```

Plot results:
```{r}
matches2 %>% 
  ggplot(aes(x = covid_days_in_hospital)) +
  geom_density(color = "blue") +
  geom_density(aes(x = ctrl_covid_days_in_hospital), color = "red") +
  labs(title = plot2title, 
       x = "Length of stay (days)")
```

View results of all tests:
```{r echo=FALSE}
ttest_output %>% 
  kable() #caption = paste(tabletitle, "-", path))
```

