---
title: "exploratory"
output: 
  html_document:
  code_folding: hide
editor_options: 
  chunk_output_type: console
---

# Setup, load packages
```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(knitr)
```

```{r}
setwd("/Users/emma/Desktop/local_only/mind_study")

adt_long <- read_csv("./adt_long.csv")
adt_long_patients <- read_csv("./adt_long_patients.csv")
covid_long <- read_csv("./covid_long.csv")
micro_long <- read_csv("./micro_long.csv")
micro_long_positives_only <- read_csv("./micro_long_positives_only.csv")
patients <- read_csv("./patients.csv")
```

# Quantify what data we have in each type of dataset by MRN:
Make venn diagram output outside of R
```{r}
# MRNs in each dataset:
adt_mrns <- adt_long %>% group_by(mrn) %>% summarize()
micro_mrns <- micro_long %>% group_by(mrn) %>% summarize()
covid_mrns <- covid_long %>% group_by(mrn) %>% summarize() 

# Compare MRNs in each dataset:
# N mrns that appear in all three datasets:
n_acm <- nrow(intersect(adt_mrns, intersect(covid_mrns, micro_mrns)))
# N mrns that appear in adt and covid but not micro datasets:
n_ac <- nrow(intersect(adt_mrns, covid_mrns)) - n_acm
# N mrns that appear in adt and micro but not covid datasets:
n_am <- nrow(intersect(adt_mrns, micro_mrns)) - n_acm
# N mrns that appear in covid and micro but not adt datasets:
n_cm <- nrow(intersect(covid_mrns, micro_mrns)) - n_acm
# N mrns that appear in adt dataset only:
n_a_only <- nrow(adt_mrns) - n_ac - n_am - n_acm
# N mrns that appear in covid dataset only:
n_c_only <- nrow(covid_mrns) - n_ac - n_cm - n_acm
# N mrns that appear in micro dataset only:
n_m_only <- nrow(micro_mrns) - n_am - n_cm - n_acm
# Total N mrns in ADT
n_a <- nrow(adt_mrns)
# Total N mrns in Covid
n_c <- nrow(covid_mrns)
# Total N mrns in micro
n_m <- nrow(micro_mrns)
```


# Exploratory analysis of mortality, etc:

assess missing deceased and hospice variables:
```{r}
#missing deceased
patients %>% 
  filter(is.na(deceased_logical)) %>% 
  select(mrn, deceased_logical, deceased_dt, 
          hospice_logical, hospice_dt) 
# Since patients includes everyone we have in either the adt, covid or micro datasets, deceased data will be missing for anyone not in the adt dataset.

# missing hospice
patients %>% 
  filter(is.na(hospice_logical)) %>% 
  select(mrn, hospice_logical, hospice_dt) 

patients %>% 
  filter(is.na(hospice_logical) & is.na(deceased_logical))
# Same number of mrns are missing both deceased and hospice outcomes.
```

Tabulate mortality among covid vs non-covid patients in adt dataset. Non-covid includes tested negative and never tested for covid
```{r}
patients %>% 
  group_by(covid_pos_ever) %>% 
  summarize(n = n(), 
            percent_deceased = mean(deceased_logical, na.rm = TRUE)*100, 
            percent_hospice = mean(hospice_logical, na.rm = TRUE)*100) %>% 
  kable()
```

Mortality by date of first event - covid patients
For this I looked at mortality among patients who EVER had covid by first event date listed
```{r}
# Plot encounters, deaths and hospice visits over time for covid pos patients:
adt_long_patients %>% 
  filter(covid_pos_ever == T) %>% # keep only covid pos patients
  group_by(mrn) %>% 
  summarize(first_adt_event_dt = min(adt_event_dt), # first event dt for each mrn
            deceased_logical= max(deceased_logical),
            hospice_logical = max(hospice_logical)) %>% 
  # one record per mrn of covid_pos_ever patients now
  mutate(first_adt_event_date = date(first_adt_event_dt)) %>% 
  group_by(first_adt_event_date) %>% 
  summarize(mortality_by_date = mean(deceased_logical, na.rm = TRUE), 
            first_encounters = n(),
            deaths = sum(deceased_logical),
            hospice = sum(hospice_logical)) %>% 
  pivot_longer(cols = c(first_encounters, deaths, hospice), names_to = "type", values_to = "n") %>% 
  ggplot(aes(x = first_adt_event_date)) +
  geom_line(aes(y = n, colour = type)) +
  ylim(0,250) +
  labs(x = "Date of first encounter", y = "N", title = "Encounters, deaths, and transfers to hospice care among covid patients by date of first encounter")

#mortality rate among covid patients agged by week 
adt_long_patients %>% 
  filter(covid_pos_ever == T) %>% # keep only covid pos patients
  group_by(mrn) %>% 
  summarize(first_adt_event_dt = min(adt_event_dt), # first event dt for each mrn
            deceased_logical= max(deceased_logical),
            hospice_logical = max(hospice_logical)) %>% 
  # one obs per mrn now
  mutate(date = date(first_adt_event_dt),
         week = week(first_adt_event_dt),
         year = year(first_adt_event_dt)) %>% 
  group_by(year, week) %>% 
  summarize(mean_date = mean(date), # should be mid week
            mortality_by_week = mean(deceased_logical, na.rm = TRUE)) %>% #aggregated mortality for the week
  ggplot(aes(x = mean_date)) +
  geom_line(aes(y = mortality_by_week)) +
  labs(x = "Week of first encounter", y = "Mortality", title = "Mortality among covid patients by week of first encounter")
```

Non-covid patients:
```{r}
# Plot encounters, deaths and hospice visits over time for covid neg patients:
adt_long_patients %>% 
  filter(covid_pos_ever == F) %>% 
  group_by(mrn) %>% 
  summarize(first_adt_event_dt = min(adt_event_dt), # first event dt for each mrn
            deceased_logical= max(deceased_logical),
            hospice_logical = max(hospice_logical)) %>% 
  # one record per mrn of covid_pos_ever patients now
  mutate(first_adt_event_date = date(first_adt_event_dt)) %>% 
  group_by(first_adt_event_date) %>% 
  summarize(first_encounters = n(),
            deaths = sum(deceased_logical, na.rm = T),
            hospice = sum(hospice_logical, na.rm = T)) %>% 
  pivot_longer(cols = c(first_encounters, deaths, hospice), names_to = "type", values_to = "n") %>% 
  ggplot(aes(x = first_adt_event_date)) +
  geom_line(aes(y = n, colour = type)) +
  ylim(0,650) +
  labs(x = "Date of first encounter", y = "N", title = "Encounters, deaths, and transfers to hospice care among non-covid patients by date of first encounter")

#mortality rate among covid patients agged by week 
adt_long_patients %>% 
  filter(covid_pos_ever == F) %>% 
  group_by(mrn) %>% 
  summarize(first_adt_event_dt = min(adt_event_dt), # first event dt for each mrn
            deceased_logical= max(deceased_logical),
            hospice_logical = max(hospice_logical)) %>% 
  # one obs per mrn now
  mutate(date = date(first_adt_event_dt),
         week = week(first_adt_event_dt),
         year = year(first_adt_event_dt)) %>% 
  group_by(year, week) %>% 
  summarize(mean_date = mean(date), # should be mid week
            mortality_by_week = mean(deceased_logical, na.rm = TRUE)) %>% #aggregated mortality for the week
  ggplot(aes(x = mean_date)) +
  geom_line(aes(y = mortality_by_week)) +
  labs(x = "Week of first encounter", y = "Mortality", title = "Mortality among non-covid patients by week of first encounter")
```

# Bacterial Infections among covid patients

Tabulate mortality among covid vs non-covid patients with vs without bacterial infection EVER:
```{r}
patients %>% 
  group_by(covid_pos_ever, bacterial_infection_ever) %>%
  summarize(n = n(), 
            percent_deceased = mean(deceased_logical, na.rm = TRUE)*100, 
            percent_hospice = mean(hospice_logical, na.rm = TRUE)*100) %>% 
  kable()
```


Plot mortality over time among EVER bacterial infection positive covid patients:
```{r}
adt_long_patients %>% 
  filter(covid_pos_ever == T,
         bacterial_infection_ever == T) %>% 
  group_by(mrn) %>% 
  summarize(first_adt_event_dt = min(adt_event_dt), # first event dt for each mrn
            deceased_logical= max(deceased_logical),
            hospice_logical = max(hospice_logical)) %>% 
  # one record per mrn of covid_pos_ever patients now
  mutate(first_adt_event_date = date(first_adt_event_dt)) %>% 
  group_by(first_adt_event_date) %>% 
  summarize(mortality_by_date = mean(deceased_logical, na.rm = TRUE), 
            first_encounters = n(),
            deaths = sum(deceased_logical),
            hospice = sum(hospice_logical)) %>% 
  pivot_longer(cols = c(first_encounters, deaths, hospice), names_to = "type", values_to = "n") %>% 
  ggplot(aes(x = first_adt_event_date)) +
  geom_line(aes(y = n, colour = type)) +
  ylim(0,45) +
  labs(x = "Date of first encounter", y = "N", title = "Encounters, deaths, and transfers to hospice care among covid patients by date of first encounter")

adt_long_patients %>% 
  filter(covid_pos_ever == T,
         bacterial_infection_ever == T) %>% 
  group_by(mrn) %>% 
  summarize(first_adt_event_dt = min(adt_event_dt), # first event dt for each mrn
            deceased_logical= max(deceased_logical),
            hospice_logical = max(hospice_logical)) %>% 
  # one obs per mrn now
  mutate(date = date(first_adt_event_dt),
         week = week(first_adt_event_dt),
         year = year(first_adt_event_dt)) %>% 
  group_by(year, week) %>% 
  summarize(mean_date = mean(date), # should be mid week
            mortality_by_week = mean(deceased_logical, na.rm = TRUE)) %>% #aggregated mortality for the week
  ggplot(aes(x = mean_date)) +
  geom_line(aes(y = mortality_by_week)) +
  labs(x = "Week of first encounter", y = "Mortality", title = "Mortality among non-covid patients by week of first encounter")
```


Look at whether bacterial infections took place before or after covid start time, and mortality for each group:
```{r}
micro_long %>% 
  filter(abnormal_result_logical == T) %>% # keep positives only
  left_join(patients) %>% 
  filter(covid_pos_ever == T) %>% # keep covid pos patients only
  mutate(bacterial_after_covid = 
           ifelse(covid_pos_dt <= order_dt, T, F), 
         bacterial_before_covid = 
           ifelse(covid_pos_dt >= order_dt, T, F) 
         ) %>% 
  group_by(mrn) %>% 
  summarize(deceased_logical = as.logical(max(deceased_logical)),
            hospice_logical = as.logical(max(hospice_logical)),
            bacterial_after_covid = as.logical(max(bacterial_after_covid)), 
            bacterial_before_covid = as.logical(max(bacterial_before_covid))) %>% 
  group_by(bacterial_after_covid, bacterial_before_covid) %>% 
  summarize(n = n(), mortality = mean(deceased_logical)) %>% 
  kable()

patients %>% 
  filter(covid_pos_ever == T,
         is.na(covid_pos_dt)) %>% 
  select(covid_pos_ever, covid_pos_dt)
  
# zI think this is keeping ppl who never got tested - check and correct!
```


~~~~~~~~~~~~~~~~ Where I stopped double checking code. may not run ~~~~~~~~~~~~~~~~~~

### Plot trajectories of covid patients who also had bacterial infections over time
Include: inpatient and outpatient, first covid positive pcr, first positive bacterial infection

First, let's see how many positive micro tests each person typically has:
```{r}
#adt_wide_covid_wide_micro_long: one observation per person per micro test.

# first, let's see if people generally have multiple positive micro tests:
adt_wide_covid_wide_micro_long %>% 
  filter(abnormal_result == "Y") %>% # now have one obs per adt event per positive micro test, only have ppl included with positive micro test
  group_by(mrn) %>% 
  summarize(n_abnormal_micro_tests = n(), ) %>% 
  ggplot(aes(x = n_abnormal_micro_tests)) +
  geom_density()
```

# Exploratory analysis for time in hospital

First, what type of visit was usually the first covid visit for patients who ever had a positive PCR?
```{r}
adt_long_covid_wide %>% 
  filter(any_pcr_pos == 1) %>% #produces 53,361 obs, one per encounter with any patient who has ever has a positive covid pcr test.
  filter(adt_visit ==1) %>% # keeps 11,312: only the first covid visit for each patient - may want to double check
  group_by(adt_care_level) %>% 
  summarize(n = n(), percent = round((n()/11312)*100, digits = 2)) %>% 
  kable()
```

Looks like almost all "first covid visits" were outpatient or emergency visits. Look into: how many were admitted afterwards in that same encounter?
```{r}
adt_long_covid_wide %>% 
  filter(any_pcr_pos == 1, # look at covid pos ever ONLY
         adt_visit ==1) %>% # consider only "first covid visits" for now. Should be one row per mrn
  mutate(admission_ordered_logical = ifelse(is.na(adt_inpatient), F, T)) %>%  # true if an admission was ordered in that same encounter
  group_by(adt_care_level) %>% 
  summarize(n = n(), n_admitted = sum(admission_ordered_logical), percent_admitted = round(n_admitted * 100 / n, digits = 2)) %>% 
  kable() %>% 
```

Look at the same as above, but consider anyone who was admitted within 7 days after or two days before  "first covid visit"
```{r}
admissions_summary <- adt_long_covid_wide %>% 
  filter(any_pcr_pos == 1) %>% # look at covid pos ever only
  group_by(mrn) %>% 
  summarize(covid_admission = as.logical(max(covid_admission)),
            deceased = as.logical(max(deceased)),
            covid_visit_start_date = date(max(covid_visits_start_date)),
            first_covid_inpatient_admission_date = min(covid_inpatient_admission_date), # look at first covid inpatient admission date
            first_covid_inpatient_discharge_date = min(covid_inpatient_discharge_date),
            date_of_death = max(death_datetime), # consider date of discharge with disposition "expired" to be date of death
            n_covid_admissions = sum(covid_admission),
            time_in_hospital = max(time_in_hospital)) # number of "covid admissions" to see if ppl ever are admitted, discharged, admitted again in short timeframe
  # mutate(first_covid_inpatient_admission_date = case_when(!is.na(first_covid_inpatient_admission_date) ~ date(first_covid_inpatient_admission_date)),
  #        first_covid_inpatient_discharge_date = case_when(!is.na(first_covid_inpatient_discharge_date) ~ date(first_covid_inpatient_discharge_date)),
  #        date_of_death = case_when(!is.na(date_of_death) ~ date(date_of_death)))

# might want to make this long anyway, so you can plot by date ^^^^

admissions_summary %>% 
  ungroup() %>% 
  summarize(n_covid_cases = n(),
            n_covid_admissions = sum(covid_admission),
            n_covid_deaths = sum(deceased, na.rm = T)) %>% 
  kable()

admissions_summary %>% 
  group_by(covid_visit_start_date) %>% 
  summarize(n_covid_cases = n()) %>% 
  ggplot(aes(x = covid_visit_start_date)) +
  geom_line(aes(y = n_covid_cases))

admissions_summary %>% 
  group_by(covid_visit_start_date) %>% 
  summarize(n_covid_inpatient_admissions = sum(covid_admission, na.rm = T)) %>% 
  ggplot(aes(x = covid_visit_start_date)) +
  geom_line(aes(y = n_covid_inpatient_admissions))

admissions_summary %>% 
  group_by(covid_visit_start_date) %>% 
  summarize(n_covid_deaths = sum(deceased, na.rm = T)) %>% 
  ggplot(aes(x = covid_visit_start_date)) +
  geom_line(aes(y = n_covid_deaths))

adm_summary_2 <- admissions_summary %>% 
  group_by(covid_visit_start_date) %>% 
  summarize(n_covid_cases = n(),
            n_covid_inpatient_admissions = sum(covid_admission, na.rm = T),
            n_covid_deaths = sum(deceased, na.rm = T),
            mean_time_in_hospital = mean(time_in_hospital, na.rm = T)) %>% 
  mutate(inpatient_admission_rate = n_covid_inpatient_admissions/n_covid_cases,
         mortality_rate_among_inpatient = n_covid_deaths/n_covid_inpatient_admissions)

adm_summary_2 %>% 
  ggplot(aes(x = covid_visit_start_date)) +
  geom_line(aes(y = inpatient_admission_rate))

adm_summary_2 %>% 
  ggplot(aes(x = covid_visit_start_date)) +
  geom_line(aes(y = mortality_rate_among_inpatient))

adm_summary_2 %>% 
  ungroup() %>% 
  ggplot(aes(x = covid_visit_start_date)) +
  geom_path(aes(y = mean_time_in_hospital))
         
```





What was the time in hospital for those who were admitted? (admitted during that initial encounter that we're calling "covid visit 1")
```{r}
# same start as previous question
adt_long_covid_wide %>% 
  filter(any_pcr_pos == 1, # look at covid pos ever ONLY
         adt_visit ==1) %>% # consider only "first covid visits" for now. Should be one row per mrn
  mutate(admission_ordered_logical = ifelse(is.na(adt_inpatient), F, T),
         adt_inpatient_date = date(adt_inpatient), # may want to eventually put all these in a separate cleaning step.
         adt_discharge_date = date(adt_discharge),
         ) %>% 
  filter(admission_ordered_logical == T) %>% 
  mutate(time_in_hospital = adt_discharge_date - adt_inpatient_date) %>% 
  ggplot(aes(y = time_in_hospital)) +
  geom_boxplot()

# Check for missingness:
adt_long_covid_wide %>% 
  filter(any_pcr_pos == 1, # look at covid pos ever ONLY
         adt_visit ==1) %>% # consider only "first covid visits" for now. Should be one row per mrn
  mutate(admission_ordered_logical = ifelse(is.na(adt_inpatient), F, T),
         adt_inpatient_date = date(adt_inpatient), # may want to eventually put all these in a separate cleaning step.
         adt_discharge_date = date(adt_discharge),
         ) %>% 
  filter(admission_ordered_logical == T) %>% 
  mutate(time_in_hospital = adt_discharge_date - adt_inpatient_date,
         is_na_time_in_hospital = is.na(time_in_hospital),
         is_na_discharge = is.na(adt_discharge)) %>% 
  summarize(nas = sum(is_na_time_in_hospital), na_percent = mean(is_na_time_in_hospital), na_discharge = sum(is_na_discharge))
# discharge date is missing for about 161 patients. 4% ish?

adt_long_covid_wide %>% 
  filter(any_pcr_pos == 1, # look at covid pos ever ONLY
         adt_visit ==1) %>% # consider only "first covid visits" for now. Should be one row per mrn
  mutate(admission_ordered_logical = ifelse(is.na(adt_inpatient), F, T),
         adt_inpatient_date = date(adt_inpatient), # may want to eventually put all these in a separate cleaning step.
         adt_discharge_date = date(adt_discharge),
         ) %>% 
  filter(admission_ordered_logical == T) %>% 
  mutate(time_in_hospital = adt_discharge_date - adt_inpatient_date,
         is_na_time_in_hospital = is.na(time_in_hospital),
         is_na_discharge = is.na(adt_discharge)) %>% 
  summarize(mean_time_in_hosp = mean(time_in_hospital, na.rm = T), 
            min_time_in_hosp = min(time_in_hospital, na.rm = T),
            p25_time_in_hosp = quantile(time_in_hospital, probs = .25, na.rm = T),
            median_time_in_hospital = median(time_in_hospital, na.rm = T),
            p75_time_in_hosp = quantile(time_in_hospital, probs = .75, na.rm = T),
            max_time_in_hosp = max(time_in_hospital, na.rm = T)) %>% 
  kable()
```

Look at time in hospital over time, using the same start from the previous code chunk:
```{r}
time_in_hospital_df <- adt_long_covid_wide %>% 
  filter(any_pcr_pos == 1, # look at covid pos ever ONLY
         adt_visit ==1) %>% # consider only "first covid visits" for now. Should be one row per mrn
  mutate(admission_ordered_logical = ifelse(is.na(adt_inpatient), F, T),
         adt_inpatient_date = date(adt_inpatient), # may want to eventually put all these in a separate cleaning step.
         adt_discharge_date = date(adt_discharge),
         ) %>% 
  filter(admission_ordered_logical == T) %>% 
  mutate(time_in_hospital = adt_discharge_date - adt_inpatient_date,
         is_na_time_in_hospital = is.na(time_in_hospital),
         is_na_discharge = is.na(adt_discharge)) %>% 
  group_by(adt_inpatient_date) %>% 
  summarize(median_time_in_hospital = median(time_in_hospital, na.rm = T),
            mean_time_in_hospital = mean(time_in_hospital, na.rm = T))

#median time in hosp by date of admission
time_in_hospital_df %>% 
  ggplot(aes(x = adt_inpatient_date)) +
  geom_line(aes(y = median_time_in_hospital)) +
  labs(title = "Median time in hospital by day of admission")

#median time in hosp by date of admission
time_in_hospital_df %>% 
  ggplot(aes(x = adt_inpatient_date)) +
  geom_line(aes(y = mean_time_in_hospital)) +
  labs(title = "Mean time in hospital by day of admission")
```

Let's do the same thing, but look at weekly:
```{r}
time_in_hosp_weekly <- adt_long_covid_wide %>% 
  filter(any_pcr_pos == 1, # look at covid pos ever ONLY
         adt_visit ==1) %>% # consider only "first covid visits" for now. Should be one row per mrn
  mutate(admission_ordered_logical = ifelse(is.na(adt_inpatient), F, T),
         adt_inpatient_date = date(adt_inpatient), # may want to eventually put all these in a separate cleaning step.
         adt_discharge_date = date(adt_discharge),
         ) %>% 
  filter(admission_ordered_logical == T) %>% 
  mutate(time_in_hospital = adt_discharge_date - adt_inpatient_date,
         is_na_time_in_hospital = is.na(time_in_hospital),
         is_na_discharge = is.na(adt_discharge),
         week_of_inpatient_admission = week(adt_inpatient_date)) %>% 
  group_by(week_of_inpatient_admission) %>% 
  summarize(median_time_in_hospital = median(time_in_hospital, na.rm = T),
            mean_time_in_hospital = mean(time_in_hospital, na.rm = T),
            week_inpatient_admission = mean(adt_inpatient_date)) 

time_in_hosp_weekly %>% 
  ggplot(aes(x = week_inpatient_admission)) +
  geom_line(aes(y = median_time_in_hospital)) +
  labs(title = "Median time in hospital by week of admission")

#mean: 
time_in_hosp_weekly %>% 
  ggplot(aes(x = week_inpatient_admission)) +
  geom_line(aes(y = mean_time_in_hospital)) +
  labs(title = "mean time in hospital by week of admission")


```

~~~~~~~~~~~~~~~~~~~~~~~~~~~~ Should run again after here ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

Add 5/10: 

First of all, do most patients have just one "covid admission"?
```{r}
adt_long_patients %>% 
  group_by(mrn) %>% 
  summarize(n_covid_admissions = sum(covid_admission)) %>% 
  ungroup() %>% 
  group_by(n_covid_admissions) %>% 
  summarize(n = n()) %>% 
  kable()
```



time in hospital violin plots for each month:
```{r}
adt_long_patients %>% 
  filter(mrn == 1000020668)

adt_long_patients %>% 
  filter(covid_admission == T) %>% # keep only "covid admissions" per our definition
  group_by(mrn) %>% 
  summarize(covid_adm_total_time_in_hosp = sum(covid_admission_time_in_hosp),
            first_covid_adm_dt = min(adt_inpatient_dt, na.rm = T)) %>% 
  ungroup() %>% 
  mutate(covid_adm_total_time_in_hosp = covid_adm_total_time_in_hosp/86400, # turn from seconds to days manually. f lubridate
         year = year(first_covid_adm_dt),
         month = month(first_covid_adm_dt),
         y_m = paste(year, month, sep = "_"),
         y_m = factor(y_m, levels = c("2020_3", "2020_4", "2020_5", "2020_6", "2020_7", "2020_8", 
                                      "2020_9", "2020_10", "2020_11", "2020_12", "2021_1", "2021_2"))) %>% 
  group_by(y_m) %>% 
  ggplot(aes(x = y_m, y = covid_adm_total_time_in_hosp)) + 
  geom_violin(aes(fill = y_m), alpha = .5) +
  labs(title = "Total time in hospital for covid admissions by date of first covid admission",
       x = "Month of first covid admission",
       y = "Days in hospital for all covid admissions",
       legend = "Month")
```

Repeat the same graph, but for individuals who survived
```{r}
adt_long_patients %>% 
  filter(covid_admission == T, # keep only "covid admissions" per our definition
         deceased_logical == F) %>% 
  group_by(mrn) %>% 
  summarize(covid_adm_total_time_in_hosp = sum(covid_admission_time_in_hosp),
            first_covid_adm_dt = min(adt_inpatient_dt, na.rm = T)) %>% 
  ungroup() %>% 
  mutate(covid_adm_total_time_in_hosp = covid_adm_total_time_in_hosp/86400, # turn from seconds to days manually. f lubridate
         year = year(first_covid_adm_dt),
         month = month(first_covid_adm_dt),
         y_m = paste(year, month, sep = "_"),
         y_m = factor(y_m, levels = c("2020_3", "2020_4", "2020_5", "2020_6", "2020_7", "2020_8", 
                                      "2020_9", "2020_10", "2020_11", "2020_12", "2021_1", "2021_2"))) %>% 
  group_by(y_m) %>% 
  ggplot(aes(x = y_m, y = covid_adm_total_time_in_hosp)) + 
  geom_violin(aes(fill = y_m), alpha = .5) +
  labs(title = "Total time in hospital for covid admissions among patients who survived",
       x = "Month of first covid admission",
       y = "Days in hospital for all covid admissions",
       legend = "Month")
```


# Micro data exploration

## General tabulations
Questions to consider: how do we define "the same infection" ? 
How do we define "end of infection"? 
If it's by a negative test following the last positive one, how do we know how soon after it needs to be to be testing for that purpose?

How many abnormal results did patients generally have?
(Note, this is just among patients in the micro dataset, ie, those who have ever had a micro test.)
```{r}
micro_long_positives_only %>% 
  group_by(mrn) %>% 
  summarize(n_abnormal_results = sum(abnormal_result_logical)) %>% 
  ungroup() %>% 
  ggplot(aes(x = n_abnormal_results)) +
  geom_histogram() 
```

How many unique organisms does each person have abnormal results for?
```{r}
micro_long_positives_only %>% 
  group_by(mrn) %>% 
  summarize(n_organisms = n_distinct(organism_name)) %>% 
  ggplot(aes(x = n_organisms)) +
  geom_histogram()

micro_long_positives_only %>% 
  group_by(mrn) %>% 
  summarize(n_abnormal_results = sum(abnormal_result_logical),
            n_organisms = n_distinct(organism_name)) %>% 
  group_by(n_organisms, n_abnormal_results) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = n_organisms, values_from = n, names_prefix = "organisms_") %>% 
  kable()
```


How many datapoints do we have for infections at different sites and species categories? 
  a: table or bar plot series disaggregated by those two things

Tabulations:
```{r}
micro_long_positives_only %>% 
  group_by(organism_name) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))

micro_long_positives_only %>% 
  group_by(organism_name, specimen_group) %>% 
  summarize(n_cases = n()) %>% 
  ungroup() %>% 
  slice_max(n_cases, n = 10) %>% 
  kable()

# blood results only: 
micro_long_positives_only %>% 
  filter(specimen_group == "Blood") %>% 
  group_by(organism_name) %>% 
  summarize(n_cases = n()) %>% 
  ungroup() %>% 
  slice_max(n_cases, n = 10) %>% 
  kable()
```

## Length of infection

For how many of these does the person have a subsequent negative test?
Do people typically have a subsequent negative test that we could use to indicate the end of the infection?
Ideas for how to answer: summarize (or mutate, actually?) to get mrn's final abnormal dt, and then see if there's another normal infection after that.
```{r}
temp <- micro_long %>% 
  filter(specimen_group == "Blood", encounter_type == "Inpatient") %>% 
  left_join(adt_long) %>% 
    # We now have one row per micro result x adt record
  filter(order_dt > adt_inpatient_dt, order_dt < adt_discharge_dt) %>% 
    # Filter so that we have one row per micro result, with only the relevant adt admission (the one that overlaps the order date and time) attached
  group_by(mrn) %>% 
  arrange(mrn, order_dt) %>% 
  select(mrn, abnormal_result_logical, organism_name, organism_category, order_dt, adt_inpatient_dt, adt_discharge_dt, adt_visit) 
    # select relevant variables for easier viewing

view(temp)
```

Try calculating end date for infections. Start with E-Coli only
to do: somewhere here, filter for only people who have ever had ecoli.
```{r}
## ecoli_mrns contains the mrns of patients who have ever had ecoli infections in blood during inpatient stays
ecoli_mrns <- unlist(
  micro_long %>% 
  filter(specimen_group == "Blood", organism_name == "ESCHERICHIA COLI", encounter_type == "Inpatient") %>% 
    mutate(mrn = as.integer(mrn)) %>% 
  select(mrn))

## calculate end of infection for e-coli infections  
temp <- micro_long %>% 
  filter(mrn %in% ecoli_mrns, 
         specimen_group == "Blood", 
         encounter_type == "Inpatient",
         organism_name == "ESCHERICHIA COLI" | is.na(organism_name)) %>% 
          # keep e-coli and negative inpatient blood results for patients who have ever had e-coli
  left_join(adt_long) %>%  # now have one row per micro result x adt_record
  filter(order_dt > adt_inpatient_dt, order_dt < adt_discharge_dt) %>% 
          # filter so that we have one row per micro result, with only the relevant adt admission data (the one that overlaps the order date and time)
          # note, this step also means that we will only have patients who are in the adt dataset included here.
  group_by(mrn, adt_visit) %>% 
          # one group per inpatient visit per patient at this point there may be visits included without any e-coli infection
  arrange(mrn, adt_visit, order_dt) %>% 
  select(mrn, abnormal_result_logical, organism_name, organism_category, order_dt, adt_inpatient_dt, adt_discharge_dt, adt_visit) %>% 
          # arrange and select relevant variables for easier viewing
  mutate(abnormal_result_dt = case_when(abnormal_result_logical == T ~ order_dt),
         negative_result_dt = case_when(abnormal_result_logical == T ~ order_dt),
         last_abnormal_result_dt = max(abnormal_result_dt, na.rm = T)) %>% # max (last) time of abnormal result
  filter(order_dt > last_abnormal_result_dt | abnormal_result_logical == T) %>% 
          # keep if abnormal result OR if it was a negative result occurring after the final abnormal result of that inpatient stay
  summarize(first_negative_result_dt = min(negative_result_dt, na.rm = T),
            first_abnormal_result_dt = min(abnormal_result_dt, na.rm = T)) %>% 
  mutate(duration_of_infection = interval(first_negative_result_dt, first_abnormal_result_dt))

  
# still some errors - need to account for NAs

view(temp)
```

Next attempt: differentiating between multiple infections based on conversation with Anne-Catrin and Jason online.
Use positives only, since they seem to use positive tests to establish duration of bacteremia 
(rather than negative tests to check for clearance). Following code just looks at 
```{r}
# make list of numbers of infections by organism and lapse times:
micro_long_positives_only %>% 
  filter(specimen_group == "Blood", encounter_type == "Inpatient") %>% 
  left_join(adt_long) %>% # We now have one row per micro result x adt record
  filter(order_dt > adt_inpatient_dt, order_dt < adt_discharge_dt) %>% 
    # Filter so that we have one row per micro result, with only the relevant adt admission 
    # (the one that overlaps the order date and time) attached
  select(mrn, abnormal_result_logical, organism_name, organism_category, order_dt, adt_inpatient_dt, adt_discharge_dt, adt_visit) %>% 
  group_by(mrn, adt_visit, organism_name) %>% 
  arrange(mrn, adt_visit, organism_name, order_dt) %>% 
  mutate(previous_order_dt = lag(order_dt, n = 1, default = NA),
         time_since_prev_ord = as.duration(order_dt - previous_order_dt),
         lapse_category = 
           case_when(time_since_prev_ord < as.duration("3 days") ~ "less than 3",
                     time_since_prev_ord < as.duration("7 days") & time_since_prev_ord > as.duration("3 days") ~ "3 to 7",
                     time_since_prev_ord < as.duration("30 days") & time_since_prev_ord > as.duration("7 days") ~ "7 to 30",
                     time_since_prev_ord > as.duration("30 days") ~ "greater than 30"),
         lapse_3_to_7 = (lapse_category == "3 to 7"),
         lapse_7_to_30 = (lapse_category == "7 to 30"),
         lapse_more_than_30 = (lapse_category == "greater than 30")) %>% # consolidate these lines later if time
  summarize(lapse_3_to_7 = max(lapse_3_to_7, na.rm = T),
            lapse_7_to_30 = max(lapse_7_to_30, na.rm = T),
            lapse_more_than_30 = max(lapse_more_than_30, na.rm = T)) %>% 
  mutate(lapse_3_to_7 = replace(lapse_3_to_7, lapse_3_to_7 == -Inf, 0),
         lapse_7_to_30 = replace(lapse_7_to_30, lapse_7_to_30 == -Inf, 0),
         lapse_more_than_30 = replace(lapse_more_than_30, lapse_more_than_30 == -Inf, 0), 
            # If the value is -Inf, it's because all values summarized in the last step were NA.
            # NA here means there was only one test, so there cannot have been a lapse btw tests.Replace -Inf with 0 (which will be false) for these cases
         lapse_3_to_7 = as.logical(lapse_3_to_7),
         lapse_7_to_30 = as.logical(lapse_7_to_30),
         lapse_more_than_30 = as.logical(lapse_more_than_30)) %>% # recode as logical
    # one line per person per inpatient stay per organism: was there a lapse in positive tests of 7-30 days? Was there a lapse of > 30 days?
  ungroup() %>% 
  group_by(organism_name) %>% 
  summarize(total_n = n(),
            n_lapse_3_to_7= sum(lapse_3_to_7),
            prop_lapse_3_to_7 = mean(lapse_3_to_7),
            n_lapse_7_to_30 = sum(lapse_7_to_30),
            prop_lapse_7_to_30 = mean(lapse_7_to_30),
            n_lapse_more_than_30 = sum(lapse_more_than_30),
            prop_lapse_more_than_30 = mean(lapse_more_than_30)) %>% 
  arrange(desc(total_n)) %>% 
  kable()
```

For now, look at just infections that do NOT have lapses of more than 30 days - I haven't quite figured out how to separate those out yet.
Will want to do something like: create infection index that assigns everything after the long lapse to one value and everything before (within that group) to another,
then group by that index variable, then summarize like I'll do below.

```{r}
# make dataset of durations of bacteremia. May want to add certain variables and make sure to keep them through summarize step at a later time.
duration_of_bacteremia <- micro_long_positives_only %>% 
  filter(specimen_group == "Blood", encounter_type == "Inpatient") %>% 
  left_join(adt_long) %>% # We now have one row per micro result x adt record
  filter(order_dt > adt_inpatient_dt, order_dt < adt_discharge_dt) %>% 
    # Filter so that we have one row per micro result, with only the relevant adt admission 
    # (the one that overlaps the order date and time) attached
  select(mrn, abnormal_result_logical, organism_name, organism_category, order_dt, adt_inpatient_dt, adt_discharge_dt, adt_visit) %>% 
  group_by(mrn, adt_visit, organism_name) %>% 
  arrange(mrn, adt_visit, organism_name, order_dt) %>% 
  mutate(previous_order_dt = lag(order_dt, n = 1, default = NA),
         time_since_prev_ord = as.duration(order_dt - previous_order_dt),
         time_since_prev_ord = replace_na(time_since_prev_ord, 0), # make 0 if NA. Should avoid warnings.
         max_lapse_in_visit = as.duration(max(time_since_prev_ord)),
         start_new_infection_dt = case_when(time_since_prev_ord > as.duration("30 days") ~ order_dt),
         start_new_infection_dt = max(start_new_infection_dt, na.rm=T),
         start_new_infection_dt = na_if(start_new_infection_dt, -Inf),
         infection_number = case_when(is.na(start_new_infection_dt) ~ 1, 
                                      # infection_number is 1 if there was never a lapse of >30 days
                                      is.na(start_new_infection_dt)==F & order_dt<start_new_infection_dt ~ 1, 
                                      # infection_number is 1 if there was a lapse of 30 days and this test took place before lapse
                                      is.na(start_new_infection_dt)==F & order_dt>=start_new_infection_dt ~ 2)) %>% 
                                      # infection_number is 2 if there was a lapse of 30 days and test took place after lapse.
  group_by(mrn, adt_visit, organism_name, infection_number) %>% # grouping by infection number allows me to separate out separate infections
  summarize(first_dt = min(order_dt, na.rm = T),
            last_dt = max(order_dt, na.rm = T),
            max_lapse_in_visit = as.duration(max(max_lapse_in_visit, na.rm = T))) %>% 
  mutate(duration_bacteremia = as.duration(last_dt - first_dt)) %>% 
  left_join(patients) %>% 
  select(mrn, adt_visit, organism_name, infection_number, covid_pos_ever, covid_pos_dt, first_dt, last_dt, max_lapse_in_visit, duration_bacteremia)

# plot density of duration of bacteremia
duration_of_bacteremia %>% 
  filter(duration_bacteremia > as.duration("0 day")) %>% # drop all the ones that have only one test.
  mutate(days_bacteremia = duration_bacteremia/86400) %>% 
  ggplot(aes(x = days_bacteremia)) +
  geom_density()

# tabulate whether bacterial infections took place before, during or after covid
duration_of_bacteremia %>% 
  ungroup() %>% 
  mutate(non_covid = (covid_pos_ever == F),
         bacterial_ended_before_covid = (covid_pos_ever == T & covid_pos_dt > last_dt),
         bacterial_started_before_covid = (covid_pos_ever == T & covid_pos_dt > first_dt & covid_pos_dt < last_dt),
         covid_started_before_bacterial = (covid_pos_ever == T & covid_pos_dt < first_dt)) %>% 
  summarize(non_covid = sum(non_covid),
            bacterial_ended_before_covid = sum(bacterial_ended_before_covid),
            bacterial_started_before_covid = sum(bacterial_started_before_covid),
            covid_started_before_bacterial = sum(covid_started_before_bacterial)) %>% 
  kable()

# when, more specifically, do bacterial infections take place with respect to covid diagnosis?
duration_of_bacteremia %>% 
  filter(covid_pos_ever == T) %>% 
  mutate(time_from_covid_to_bacterial = (first_dt - covid_pos_dt)/86400) %>% 
    # time from covid to bacterial is the duration of time from first positive covid pcr test
    # until first positive test for bacterial infection for the given infection.
    # negative times indicate that bacterial infection started (perhaps ended) before covid started
  ggplot(aes(x = time_from_covid_to_bacterial)) +
  geom_histogram() +
  labs(x = "days elapsed from first positive covid pcr to first positive test for this infection",
       title = "time from covid to bacterial infection")

#How many of these bacterial infections occur during same hospitalization as covid infection?:
duration_of_bacteremia %>% 
  mutate(covid_and_bacterial_same_admission = (covid_pos_dt >= inpatient))

***********************************************

# plot duration of bacteremia for infections among non-covid patients:
duration_of_bacteremia %>% 
  filter(non_covid==T) %>% 
  filter(duration_bacteremia > as.duration("0 day")) %>% # drop all the ones that have only one test.
  mutate(days_bacteremia = duration_bacteremia/86400) %>% 
  ggplot(aes(x = days_bacteremia)) +
  geom_density()

duration_of_bacteremia %>% 
  filter(covid_started_before_bacterial==T) %>% 
  filter(max_lapse_in_visit < as.duration("7 days"),
         duration_bacteremia > as.duration("0 day")) %>% # drop all the ones that have only one test.
  mutate(days_bacteremia = duration_bacteremia/86400) %>% 
  ggplot(aes(x = days_bacteremia)) +
  geom_density()

# Actually look at most common blood infections:
duration_of_bacteremia %>% 
  filter()
  group_by(organism_name) %>% 
  summarize(n_cases = n()) %>% 
  ungroup() %>% 
  slice_max(n_cases, n = 10) %>% 
  kable()
  

## will want to edi thte below to be for all organisms, not just staph epi

for(i in c("STAPHYLOCOCCUS EPIDERMIDIS",
           "ESCHERICHIA COLI", 
           "METHICILLIN-SUSCEPTIBLE STAPHYLOCOCCUS AUREUS", 
           "KLEBSIELLA PNEUMONIAE", 
           "STAPHYLOCOCCUS HOMINIS",
           "METHICILLIN-RESISTANT STAPHYLOCOCCUS AUREUS",
           "PSEUDOMONAS AERUGINOSA",
           "STAPHYLOCOCCUS CAPITIS",
           "ENTEROCOCCUS FAECALIS",
           "VANCOMYCIN-RESISTANT ENTEROCOCCUS FAECIUM",
           "CANDIDA GLABRATA")) {
  
  n <- nrow(duration_of_bacteremia %>% 
                filter(max_lapse_in_visit < as.duration("7 days"),
                       duration_bacteremia > as.duration("0 seconds"),
                       # drop all the ones that have only one test.
                organism_name == i))
  
  plot <- duration_of_bacteremia %>% 
    filter(max_lapse_in_visit < as.duration("7 days"),
           duration_bacteremia > as.duration("0 seconds"), # drop all the ones that have only one test.
           organism_name == i) %>% # filter by organism name 
    mutate(days_bacteremia = duration_bacteremia/86400) %>% 
    ggplot(aes(x = days_bacteremia)) +
    geom_density() +
    labs(title = paste("Duration of bacteremia for ", i, "; N = ", n, sep = ""))
  
  print(plot)
  
  n2 <- nrow(duration_of_bacteremia %>% 
                filter(max_lapse_in_visit < as.duration("7 days"),
                       duration_bacteremia > as.duration("3 days"),
                       # drop all the ones that have only one test.
                organism_name == i))
  
  plot2 <- duration_of_bacteremia %>% 
    filter(max_lapse_in_visit < as.duration("7 days"),
           duration_bacteremia > as.duration("3 days"), # drop all the ones that have only one test.
           organism_name == i) %>% # filter by organism name 
    mutate(days_bacteremia = duration_bacteremia/86400) %>% 
    ggplot(aes(x = days_bacteremia)) +
    geom_density() +
    labs(title = paste("Duration of bacteremia for Staphylococcus Epidermidis - patients with 3 days + only; N = ", n2, sep = ""))
  
  print(plot2)
  
   n3 <- nrow(duration_of_bacteremia %>% 
                filter(max_lapse_in_visit < as.duration("30 days"),
                       duration_bacteremia > as.duration("3 days"),
                       # drop all the ones that have only one test.
                organism_name == i))
  
  plot3 <- duration_of_bacteremia %>% 
    filter(max_lapse_in_visit < as.duration("30 days"),
           duration_bacteremia > as.duration("3 days"), # drop all the ones that have only one test.
           organism_name == i) %>% # filter by organism name 
    mutate(days_bacteremia = duration_bacteremia/86400) %>% 
    ggplot(aes(x = days_bacteremia)) +
    geom_density() +
    labs(title = paste("Duration of bacteremia for Staphylococcus Epidermidis - patients with 3 days + only; N = ", n3, sep = ""))
  
  print(plot3)
           }


```












admission trajectories for covid AND bacterial infection patients:

From Sasi:

### Detection date plot for Candida
temp <- d[org.group == 'Candida', j = list(min.date = min(spec.date), max.date = max(spec.date)), by = list(is.covid, mrn, spec.group)]
temp <- orderBy(~min.date, temp)
temp$mrn = factor(temp$mrn, levels = unique(temp$mrn))
temp$spec.group = factor(temp$spec.group, levels = c('Urine', 'Blood', 'Respiratory', 'Other'))
ggplot(temp) +
  geom_errorbarh(aes(y = factor(mrn), xmin = min.date, xmax = max.date, color = is.covid)) +
  geom_point(data = d[org.group == 'Candida', j = list(spec.date, is.covid, mrn)],
             aes(y = factor(mrn), x = spec.date, color = is.covid), alpha = .5, size= .9) +
  labs(y = 'Subject id', x = 'Specimen date', title = 'Candida Specimen') +
  theme_minimal() + theme(axis.text.y = element_blank(), legend.position = 'bottom', panel.grid.major.y =  element_blank())
  
```{r}
micro_long %>% 
  filter(abnormal_result_logical == T,
         organism_name == "METHICILLIN-SUSCEPTIBLE STAPHYLOCOCCUS AUREUS") %>% 
  ggplot(aes(x = order_dt)) +
  geom_histogram()
```

  
  
  