---
title: "exploratory"
output: 
  html_document:
  code_folding: hide
editor_options: 
  chunk_output_type: console
---

# Setup, load packages
```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(knitr)
```

# Load datasets, data cleaning, merges

## Load datasets
Set working directory to desktop -> local_only -> mind_study. 
I'm keeping data there so that it does not sync to dropbox, and is stored locally only instead.
All code files will be kept in dropbox directory.
```{r}
setwd("/Users/emma/Desktop/local_only/mind_study")

#ADT wide:
adt_wide_orig <- read_csv("./wide_files_2021_08_25/adt_summary2021-08-25.csv")

#ADT long:
adt_long_orig <- read_csv("./long_files_2021_08_25/adt_long2021-08-25.csv")

#Covid wide:
covid_wide_orig <- read_csv("./wide_files_2021_08_25/covid_testing_wide2021-08-25.csv")

#Covid long:
covid_long_orig <- read_csv("./long_files_2021_08_25/covid_pcr_long2021-08-25.csv")

#Micro long:
micro_long_orig <- read_csv("./long_files_2021_08_25/micro_long2021-08-25.csv")

#Problems wide:
problems_wide_orig <- read_csv("./wide_files_2021_08_25/problem_summary2021-08-25.csv")

#Problems long:
problems_long_orig <- read_csv("./long_files_2021_08_25/problem_long2021-08-25.csv")
```

## Data cleaning
Use all long datasets, do basic cleaning and tidying steps first, then merge later if needed for analyses and viz. 
One dataset per type of observation unit. 

### adt_long
Unit of observation: adt encounter. Includes admission, discharge, and transfer data
```{r}
adt_long <- adt_long_orig %>% 
  janitor::clean_names() %>% 
  select(-x1) %>% 
  rename(covid_visits_start_dt = covid_visits_start_date,
         adt_event_dt = adt_event_time,
         adt_ed_admission_dt = adt_ed_admission,
         adt_inpatient_dt = adt_inpatient,
         adt_discharge_dt = adt_discharge) %>%  # first pos pcr test or first neg pcr test if never pos in datetime format
  mutate(covid_visits_start_date = date(covid_visits_start_dt), # in date format
         adt_event_date = date(adt_event_dt),
         adt_ed_admission_date = date(adt_ed_admission_dt),
         adt_inpatient_date = date(adt_inpatient_dt),
         adt_discharge_date = date(adt_discharge_dt)) %>% 
  filter(adt_event_dt < ymd(20210301)) 
# this dataset was updated in Feb 2021, so dates later than march 2021 are errors.
# not sure how to keep this up to date once it's later in the year and these are indistinguishable from correct data.
```


### covid_long
Unit of observation: pcr test

Check: should we use covid_pcr_result or covid_pcr_result2 as the true/final result?:
```{r}
covid_long_orig %>% 
  filter(covid_pcr_result != covid_pcr_result2) %>% 
  select(covid_pcr_result, covid_pcr_result2)
```
All observations where covid_pcr_result differs from covid_pcr_result2 have
covid_pcr_reult = indeterminate and covid_pcr_result2 = positive. Use covid_pcr_result2 as the true/final result

Check: does any_pcr_pos match covid_pcr_result or covid_pcr_result2?
```{r}
#check any_pcr_pos variable
covid_wide_orig %>% 
  mutate(covid_pos = ifelse(is.na(any_pcr_pos), FALSE, TRUE)) %>% 
  summarize(sum_covid_pos = sum(covid_pos))

#check covid results variables
covid_long_orig %>% 
  mutate(covid_pcr_result1_logical = (covid_pcr_result == "Positive"),
         covid_pcr_result2_logical = (covid_pcr_result2 == "Positive")) %>% 
  group_by(mrn) %>% 
  summarize(covid_pcr_result1 = max(covid_pcr_result1_logical),
            covid_pcr_result2 = max(covid_pcr_result2_logical)) %>% 
  ungroup() %>% 
  summarize(covid_pcr_result1 = sum(covid_pcr_result1, na.rm = T), 
            covid_pcr_result2 = sum(covid_pcr_result2, na.rm = T))

#covid_pcr_result1 and covid_pcr_result2 now seem to match
```
PCR result 1 and pcr result 2 now seem to match (5/26/21)

```{r}
covid_long <- covid_long_orig %>% 
  janitor::clean_names() %>% 
  select(mrn, covid_pcr_date, covid_pcr_accession:machine) %>% 
  rename(covid_pcr_dt = covid_pcr_date) %>% 
  mutate(covid_pcr_date = date(covid_pcr_dt),
         covid_pos_logic = (covid_pcr_result2 == "Positive"))
```


### micro_long
Unit of observation: micro test
```{r}
micro_long <- micro_long_orig %>% 
  janitor::clean_names() %>% 
  select(mrn:organism_group) %>% 
  rename(collected_dt = collected_date_time,
         order_dt = order_date_time) %>% # put date time variables in consistent naming format.
  mutate(abnormal_result_logical = case_when(is.na(abnormal_result) ~ F,
                                             abnormal_result == "Y" ~ T))
```

### patients
Unit of observation: patient. 
This dataset includes individuals for whom we have covid test information. 
Some of these people, but not all, are included in the adt dataset. All mrns in the 
adt dataset are included in the covid data, and therefore in the patients_long df created here. 
Pull patient-level data from the covid dataset, and a few other patient-level variables from the other datasets.

Additional basic info to consider for patients:
 

Check: which MRNs in the covid testing data are also in the adt data, and vice versa?
```{r}
# How many of the mrns in the covid_wide data are also in the adt_wide data?
sum(covid_wide_orig$mrn %in% adt_wide_orig$mrn)
# 88315 out of 120031 obs in the covid_wide_orig dataset are in the adt_wide_orig dataset.
# all mrns in the adt data are in the covid data.
```

Set up a few other patient-level variables to include:
```{r}
# deceased or hospice outcome
temp_deceased_hospice <- adt_long %>% 
  mutate(deceased_logical = adt_discharge_disp == "Expired",
         hospice_logical = adt_discharge_disp == "Hospice/Medical Facility",
         deceased_dt = case_when(deceased_logical == T ~ adt_discharge_dt),
         hospice_dt = case_when(hospice_logical == T ~ adt_discharge_dt)) %>% 
  group_by(mrn) %>% 
  summarize(deceased_logical = as.logical(max(deceased_logical)),
            hospice_logical = as.logical(max(hospice_logical)),
            deceased_dt = max(deceased_dt),
            hospice_dt = max(hospice_dt)) %>% 
  mutate(adt_dataset = T)

#checkpoint:
temp_deceased_hospice %>% filter(is.na(deceased_logical))

# covid positive ever
temp_covid_pos <- covid_long %>% 
  group_by(mrn) %>% 
  mutate(covid_pos_dt = case_when(covid_pos_logic == T ~ covid_pcr_dt)) %>% 
  summarize(covid_pos_ever = as.logical(max(covid_pos_logic)),
            covid_pos_dt = min(covid_pos_dt, na.rm = T)) %>% 
  mutate(covid_pos_dt = na_if(covid_pos_dt, Inf))
    
# checkpoint: covid_pos_dt should be NA if covid_pos_ever is F, and should not be NA otherwise
temp_covid_pos %>% filter(covid_pos_ever == T & is.na(covid_pos_dt))
temp_covid_pos %>% filter(covid_pos_ever == F & is.na(covid_pos_dt) == F)

# micro bacterial infection ever
temp_micro <- micro_long %>% 
  group_by(mrn) %>%
  summarize(bacterial_infection_ever = max(abnormal_result_logical)) %>% 
  mutate(micro_dataset = T)

# bmi
temp_bmi <- covid_wide_orig %>% 
  select(mrn, bmi, bmi_date, bmi_source)

# problems. We're interested in: kidney disease, cardiovascular disease, diabetes
# cdk = chronic kidney disese, dm = diabetes mellitus, htn = hypertension, cad = coronary artery disease, 
temp_problems <- problems_wide_orig %>% 
  select(mrn, dx_ckd, ckd_date, dx_dm, dm_date, dx_htn, htn_date, dx_cad, dx_cad_date, dx_hrt_failure, hrt_failure_date)

# checkpoint: any duplicates in temp_problems?
temp_problems %>% count(mrn) %>% count()
```

Create patients dataframe.
Patients should include all mrns that we have.
adt_dataset, covid_dataset and micro_dataset variables indicate whether the mrn appears in the corresponding dataset.
```{r}
patients  <- covid_wide_orig %>% 
  janitor::clean_names() %>% 
  select(mrn, dob, sex, zipcode, 
         race, ethnicity, combined_race_ethnicity, 
         insurance_type, insurance_plan) %>% 
  mutate(covid_dataset = T) %>% 
  full_join(temp_deceased_hospice, by = "mrn") %>% 
  full_join(temp_covid_pos, by = "mrn") %>%
  full_join(temp_micro, by = "mrn") %>% 
  replace_na(list(adt_dataset = F, covid_dataset = F, micro_dataset = F)) %>% 
  left_join(temp_bmi) %>% 
  full_join(temp_problems)
  
rm(temp_deceased_hospice, temp_covid_pos, temp_micro, temp_bmi) # don't need this anymore
```


## Merges and subsets

### Merge: adt_long_patients
One row per adt per patient
```{r}
adt_long_patients <- left_join(adt_long, patients, by = "mrn")
```

### Troubleshoot: discrepencies btw adt_wide and adt_long datasets:
```{r}
# mrns that are in adt_wide:
adt_wide_mrns <- adt_wide_orig %>% select(mrn)
# mrns that are in adt_long:
adt_long_mrns <- adt_long_orig %>% group_by(mrn) %>% summarize() %>% select(mrn)

# mrns that appear in adt_long but not adt_wide:
mrns_adt_long_not_wide <- setdiff(adt_long_mrns, adt_wide_mrns)
mrns_adt_long_not_wide
nrow(mrns_adt_long_not_wide)/nrow(adt_long_mrns)
# about 3% of mrns in adt_long don't appear in adt_wide

#mrns that appear in adt_wide but not adt_long:
mrns_adt_wide_not_long <- setdiff(adt_wide_mrns, adt_long_mrns)
nrow(mrns_adt_wide_not_long)/nrow(adt_wide_mrns)

# look at records of patients in adt_long but not adt_wide:
adt_long_not_wide <- left_join(mrns_adt_long_not_wide, adt_long, by = "mrn")
adt_long_not_wide %>% 
  summarize(min_dat = min(adt_event_date, na.rm = T), 
            median_date = median(adt_event_date, na.rm = T),
            max_date = max(adt_event_date, na.rm = T)
           )
# no clear date pattern on who was included in adt_wide and who was not...
```

### Merge: adt_long_micro_long_patients
Should contain one row per admission per micro test per patient.
eg, patient A has 3 adt observations (ever) * 12 micro tests (ever) = 36 observations for patient A in this dataset.
```{r}
adt_long_micro_long_patients <- adt_long %>% 
  left_join(micro_long, by = "mrn") %>% 
  left_join(patients, by = "mrn")
```


# Data Cleaning

## adt_long_patients


Add covid admission and time in hospital variables
```{r}
adt_long_patients <- adt_long_patients %>% 
  mutate(dur_covid_to_inpatient = 
           case_when(!is.na(adt_inpatient_dt) & covid_pos_ever == T ~ adt_inpatient_dt - covid_pos_dt),
         # dur_covid_to_admission is the time interval between the datetime of first positive PCR test and the inpatient admission datetime, for patients who ever had pos pcr test.
        covid_admission = case_when(dur_covid_to_inpatient < duration(day = 7) & dur_covid_to_inpatient > duration(day = -2) ~ T, 
                                    dur_covid_to_inpatient > duration(day = 7) ~ F,
                                    dur_covid_to_inpatient < duration(day = -2) ~ F,
                                    is.na(dur_covid_to_inpatient) ~ F),
        # covid_admission is true if the admission took place within 7 days after or 2 days before time of first positive PCR test.
        covid_inpatient_admission_dt = ifelse(covid_admission == T, adt_inpatient_dt, NA),
        # datetime of inpatient admission if it was a "covid admission"
        covid_inpatient_discharge_dt = ifelse(covid_admission == T, adt_discharge_dt, NA),
        #datetime of discharge if it was a "covid admission"
        covid_admission_time_in_hosp = case_when(covid_admission == T ~ as.duration(adt_discharge_dt - adt_inpatient_dt)))
        # covid_admission_time_in_hospital is the duration btw admission and discharge for a covid admission. 
```


temp df will define more specific version of covid admission variable. This is taken from mind_exploratory 10/18/21. Once set up, this can be merged in with adt_long_patient
```{r}
temp <- adt_long_patients %>% 
  group_by(mrn) %>% 
  mutate(covid_first_admission_ever = max (covid_admission)) %>% 
    # covid_first_admission is an indicator of whether patient ever had an admission considered a covid admission by our previous criteria.
  filter(covid_first_admission_ever == T, 
         dur_covid_to_inpatient > -(as.duration("2 days"))) %>% 
    # keep only patients who ever had a covid admission per our previous criteria and visits taking place later than two-days-before time of covid pos test.
  arrange(mrn, adt_inpatient_dt) %>% 
  mutate(previous_discharge_dt = lag(adt_discharge_dt, n = 1, default = NA), 
            # create var with time of previous discharge
         time_since_prev_discharge = as.duration(adt_inpatient_dt - previous_discharge_dt),
            # create var with time between previous discharge and this line's admission to inpatient care
         time_since_prev_discharge = replace_na(time_since_prev_discharge, 0), 
             # make 0 if NA, should avoid warnings
         days_since_prev_discharge = time_since_prev_discharge/86400,
         start_non_covid_admissions_dt = case_when(time_since_prev_discharge > as.duration("2 days") & dur_covid_to_inpatient > as.duration("7 days") ~ adt_inpatient_dt),
            # start date of admissions that are not considered to be covid admissions, because > 2 days have elapsed since previous visit, AND visit starts after -2-7 day window.
         start_non_covid_admissions_dt = min(start_non_covid_admissions_dt, na.rm=T),
         start_non_covid_admissions_dt = na_if(start_non_covid_admissions_dt, -Inf),
             # Establish start of non-covid admissions as inpatient admission dt if there has been a lapse of > 48 hours
         after_lapse = (adt_inpatient_dt >= start_non_covid_admissions_dt),
             # Is the inpatient time later than the start of non-covid admissions?
         covid_admission2 = if_else(covid_admission == T, T, if_else(adt_inpatient_dt < start_non_covid_admissions_dt, T, F))) %>% 
             # covid_admission2 will be true if: 1) if covid_admission was true (if admission starts within -2 - 7 days after covid test) OR 2) if admission starts after this but with lapses of < 48 hours since a previous appointment. (if someone keeps leaving and being re-admitted in < 48 hrs, all visits will be considered to be covid visits until they have been out of the hospital for at least 48 hrs.) else, false.
  select(encounter_id, covid_admission2)
```

Merge temp with adt_long_patients: This will add covid_admission2 variable.
```{r}
adt_long_patients <- adt_long_patients %>% 
  left_join(temp)
```


Checkpoint: There should be a relatively small group of observations with covid_admission2 == T and covid_admission == F. 
```{r}
adt_long_patients %>% 
  filter(covid_admission == F, covid_admission2 == T) %>% 
  count()

adt_long_patients %>% 
  filter(covid_admission == T, covid_admission2 == F) %>% 
  count()
```

Make mini dataset from micro_long data:
- one mrn per patient who has ever had an abnormal result for a bacterial infection
- type of culture
- type of pathogen?
- Date of first abnormal result, if so. NA if not. 
```{r}
micro_long_positives_only <- micro_long %>% 
  filter(abnormal_result_logical == T) #%>% # keep only if it was an abnormal result
  #group_by(mrn) %>% 
  #summarize(first_abnormal_result = min(order_dt))
```

## Bacterial infections

This is the dataset I originally called "duration of bacteremia". It contains one observation per infection for blbacterial infections from the micro dataset.


```{r}
# make dataset of blood bacterial infections:
bacterial_infections <- micro_long_positives_only %>% 
  filter(specimen_group == "Blood") %>% 
  left_join(adt_long) %>% # We now have one row per micro result x adt record
  mutate(collected_or_order_dt = pmin(collected_dt, order_dt, na.rm = T)) %>% 
    # Some of the records have order_dt, others have collected_dt. This should give us the one that's non-missing or the earlier of the two.
  filter(collected_or_order_dt > adt_inpatient_dt, collected_or_order_dt < adt_discharge_dt,
         !(is.na(adt_inpatient_dt))) %>% 
    # Filter so that we have one row per micro result, with only the relevant adt admission 
    # (the one that overlaps the order date and time) attached
    # Filter adt_inpatient_dt is not NA, to ensure it is inpatient admissions only.
  select(mrn, abnormal_result_logical, organism_name, organism_category, collected_or_order_dt, collected_dt, order_dt, adt_inpatient_dt, adt_discharge_dt, adt_visit) %>% 
  group_by(mrn, adt_visit, organism_name) %>% 
  arrange(mrn, adt_visit, organism_name, collected_or_order_dt) %>% 
  mutate(ccollected_or_order_dt = as_datetime(collected_or_order_dt),
         previous_collected_or_order_dt = lag(collected_or_order_dt, n = 1, default = NA),
         time_since_prev = as.duration(collected_or_order_dt - previous_collected_or_order_dt),
         time_since_prev = replace_na(time_since_prev, 0), # make 0 if NA. Should avoid warnings.
         max_lapse_in_visit = as.duration(max(time_since_prev)),
         start_new_infection_dt = case_when(time_since_prev > as.duration("30 days") ~ collected_or_order_dt),
         start_new_infection_dt = max(start_new_infection_dt, na.rm=T),
         start_new_infection_dt = na_if(start_new_infection_dt, -Inf),
         infection_number = case_when(is.na(start_new_infection_dt) ~ 1, 
                                      # infection_number is 1 if there was never a lapse of >30 days
                                      is.na(start_new_infection_dt)==F & collected_or_order_dt < start_new_infection_dt ~ 1, 
                                      # infection_number is 1 if there was a lapse of 30 days and this test took place before lapse
                                      is.na(start_new_infection_dt)==F & collected_or_order_dt >= start_new_infection_dt ~ 2)) %>% 
                                      # infection_number is 2 if there was a lapse of 30 days and test took place after lapse.
  group_by(mrn, adt_visit, organism_name, infection_number) %>% # grouping by infection number allows me to separate out separate infections
  summarize(first_dt = min(collected_or_order_dt, na.rm = T),
            last_dt = max(collected_or_order_dt, na.rm = T),
            max_lapse_in_visit = as.duration(max(max_lapse_in_visit, na.rm = T))) %>% 
  mutate(duration_bacteremia = as.duration(last_dt - first_dt)) %>% 
  left_join(patients) %>% 
  left_join(adt_long) %>% # We now have one row per micro result x adt record
  filter(first_dt > adt_inpatient_dt, first_dt < adt_discharge_dt) %>% 
    # repeat this step from the beginning so we have all the data from patients and adt available.
    mutate(covid_and_bacterial_same_admission = (covid_pos_dt >= (adt_event_dt - as.duration("48 hours")) & covid_pos_dt <= adt_discharge_dt),
           covid_and_bacterial_same_admission = replace_na(covid_and_bacterial_same_admission, F))
```




adt_patients_bacterial_temp add 7/5-7/6: Set up dataset of covid stays merged with bacterial infections
```{r}
#first, set up bacterial infections dataset:
bacterial_infections_temp <- bacterial_infections %>% 
  select(mrn, adt_visit, organism_name, first_dt, last_dt, max_lapse_in_visit, duration_bacteremia,
         adt_inpatient_dt, covid_and_bacterial_same_admission) %>% 
  filter(organism_name!="STAPHYLOCOCCUS EPIDERMIDIS" & organism_name!="STAPHYLOCOCCUS HOMINIS" & organism_name!="STAPHYLOCOCCUS CAPITIS")
    # leave out these common contaminants, may want to refine this later.

covid_admissions_bacterial <- adt_long_patients %>% 
  filter(covid_admission2 == T) %>% # keep only "covid admissions" based on the covid_admissions2 definition.
  distinct() %>% # remove duplicates! Like for that one person with enterobacter somthing
  left_join(bacterial_infections_temp, by = c("mrn" = "mrn", "adt_inpatient_dt" = "adt_inpatient_dt")) %>% 
    # join by mrn and by inpatient admission, so each should be matched only with bacterial infections taking place during that inpatient admission.
    mutate(bacterial_infection = case_when(is.na(organism_name) ~ F, is.na(organism_name) == F ~ T))
    # bacterial infection is true if the observation has a bacterial infection, false otherwise.
    # kind of the same as covid_and_bacterial_same_admission, may want to simplify later.
```

Bacterial_infections_urine" Repeat dataset of bacterial infections, set up exact same way as above, except with urine cultures only. Only includes infections during inpatient stays, just like above.


# Write each dataframe to csv
To be loaded and used in the exploratory file:
```{r}
setwd("/Users/emma/Desktop/local_only/mind_study")

write_csv(adt_long, "./adt_long.csv")
write_csv(adt_long_patients, "./adt_long_patients.csv")
write_csv(covid_long, "./covid_long.csv")
write_csv(micro_long, "./micro_long.csv")
write_csv(micro_long_positives_only, "./micro_long_positives_only.csv")
write_csv(patients, "./patients.csv")
write_csv(bacterial_infections, "./bacterial_infections.csv")
write_csv(bacterial_infections_urine, "./bacterial_infections_urine.csv")
write_csv(bacterial_infections_resp, "./bacterial_infections_resp.csv")
write_csv(covid_admissions_bacterial, "./covid_admissions_bacterial.csv")

```

